Sperm should evolve to make female meiosis fair.

### 2014-12-22 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
updated the main paper

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
	creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
\yb{INSERT BIOLOGY HERE}
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this 
	apparent weakness in the defense against meiotic drive. \newline

@@

Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-s_h$)
	further constrains the evolution of a self-promoting driver. 
In fact, with any heterozygous fitness cost, a rare self-promoting
	driver is always selected against. 
However, this case also displays bistability -- 
	when $s$ is sufficiently small (Equation \ref{sForHetFix}) this allele fixes deterministically if its 
	frequency exceeds some threshold  (Equation \ref{thresh2}, exact results in 
	Figure \ref{bistable_additive}).
This bistability prevents self-promoting drivers from invading 	
	reasonably sized populations, and assures that if they do invade, they will rapidly fix.
Our model therefore predicts that self-promoting drivers will not be
	observed as stable polymorphisms in natural populations. 
This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
	precludes the evolution of an
	allele that suppresses this form of meiotic drive in females. \newline

@@

\yb{Our model of a sperm-acting drive suppressor unlinked to a female driver (Model 6) also predicts that sperm should evolve to prevent meiotic drive; 
	however, it contains no simple mechanism to maintain polymorphism for sperm-acting drive suppression.
Given the benefit to sperm of hampering female drive, 
	drive-suppressing sperm are often likely to be fixed within a species, 
	making the hypothesis of sperm-acting drive-suppression difficult to test from intra-population crosses. 
However, crosses between populations or species are likely to provide critical tests of our theory -- 
	specifically we predict that female meiosis will be less fair when a species (or population) is 
	fertilized by heterospecific sperm because either such sperm have not evolved to counter novel female meiotic drivers, 
	or because antagonistic coevolution between a driver-suppressor pair has been independent since two populations have separated. 
We can therefore predict that segregation in F1 females backcrossed to parental species will likely be biased, 
	with a deficit of transmission of the paternal species allele from the F1  female.   
These predictions follow straightforwardly from the
	theory presented above; however, we caution that tests of meiotic drive, 
	and especially sperm-dependent meiotic drive require a high standard
	of evidence  to exclude plausible alternative hypotheses such as genotypic 
	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
}

@@

Our results also suggests that phylogenetic variation in the stage of female meiosis when
        fertilization occurs may influence the prevalence of  female meiotic drive.
For example, centromeric drive may be more common  in taxa where females complete MI before fertilization, 
	as compared to species in which sperm interact with eggs arrested  in MI, 
	because in the prior case, sperm-based modifiers can only intercede during the second, but not the first  meiotic division. 
As a potential test of this hypothesis, the speed of centromere
	turnover could be compared in species in which sperm interact with eggs paused at MI and MII 
	(assuming the pace of centromere turnover serves as a proxy for the frequency of MI drivers). 
\yb{INSERT BIOLOGY HERE}
\newline

\section*{Conclusion}
Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[for a review and evaluation of such
  hypotheses see][]{Mira1998}, in fact, we presuppose its existence.
However, given the existence of female meiotic arrest, 
	and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09},
	the nature of the meiotic arrest and interactions between
        sperm and egg may  be molded by selection to reduce the opportunity for female meiotic drive, 
	and counteracted by selfish drivers evolving to overcome these adaptations.

>>>>

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only
after fertilization \citep{Masui_book}, creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
\gc{The timing of enty of sperm into the egg relative to the
  progression of female meiosis is variable across animals, as is the
  stage at which female meiosis
  stalls without fertilization (we show the phylogenetic distribution
  in supplementary Figure XXX).}
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this 
	apparent weakness in the defense against meiotic drive. \newline

@@

Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-s_h$)
	further constrains the evolution of a self-promoting driver. 
In fact, with any heterozygous fitness cost, a rare self-promoting
	driver is always selected against. 
However, this case also displays bistability -- 
	when $s$ is sufficiently small (Equation \ref{sForHetFix}) this allele fixes deterministically if its 
	frequency exceeds some threshold  (Equation \ref{thresh2}, exact results in 
	Figure \ref{bistable_additive}).
This bistability prevents self-promoting drivers from invading 	
	reasonably sized populations, and assures that if they do invade, they will rapidly fix.
Our model therefore predicts that self-promoting drivers will not be
	observed as stable polymorphisms in natural populations. 
This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
	precludes the evolution of an
	allele that suppresses this form of meiotic drive in females. 
\gc{Allowing for arbitrary levels of inbreeding (in the form of
  self-fertilization), as opposed to the panmictic model explored
  above -- more thoroughly aligns the interests of both parents and
  parental chromosomes, restricting further the possibility for
  invasion of both traditional female drivers and 'self-promoting'
  drivers (Figures SXX-SXY). Additionally, because inbreeding reduces
  the frequency of heterozygotes, the invasion and fixation criteria
  converge as both become more strict with selfing promoting rates.} \newline

@@

\yb{Our model of a sperm-acting drive suppressor unlinked to a female driver (Model 6) also predicts that sperm should evolve to prevent meiotic drive; 
	however, it contains no simple mechanism to maintain polymorphism for sperm-acting drive suppression.
Given the benefit to sperm of hampering female drive, 
	drive-suppressing sperm are often likely to be fixed within a species, 
	making the hypothesis of sperm-acting drive-suppression difficult to test from intra-population crosses. 
However, crosses between populations or species are likely to provide critical tests of our theory -- 
	specifically we predict that female meiosis will be less fair when a species (or population) is 
	fertilized by heterospecific sperm because either such sperm have not evolved to counter novel female meiotic drivers, 
	or because antagonistic coevolution between a driver-suppressor pair has been independent since two populations have separated. 
We can therefore predict that segregation in F1 females backcrossed to parental species will likely be biased, 
	with a deficit of transmission of the paternal species allele from the F1  female.   
These predictions follow straightforwardly from the
	theory presented above; however, we caution that tests of meiotic drive, 
	and especially sperm-dependent meiotic drive require a high standard
	of evidence  to exclude plausible alternative hypotheses such as genotypic 
	inviability including epistatic maternal by zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
}

@@

Our results also suggests that phylogenetic variation in the stage of female meiosis when
        fertilization occurs \gc{(see Supplementary Figure XXX)} may influence the prevalence of  female meiotic drive.
For example, centromeric drive may be more common  in taxa where females complete MI before fertilization, 
	as compared to species in which sperm interact with eggs arrested  in MI, 
	because in the prior case, sperm-based modifiers can only intercede during the second, but not the first  meiotic division. 
As a potential test of this hypothesis, the speed of centromere
	turnover could be compared in species in which sperm interact with eggs paused at MI and MII 
	(assuming the pace of centromere turnover serves as a proxy for the frequency of MI drivers). 
\newline

\section*{Conclusion}
Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[for a review and evaluation of such
  hypotheses see][]{Mira1998}, in fact, we presuppose its existence.
However, given the existence of female meiotic arrest, 
	and that its timing and mechanistic details are variable
        across species \citep[Supplementary Figure XXX, and][ ]{Masui_book,Karrbook:09},
	the nature of the meiotic arrest and interactions between
        sperm and egg may  be molded by selection to reduce the opportunity for female meiotic drive, 
	and counteracted by selfish drivers evolving to overcome these adaptations.
